---
title: "Endometrial cancer study"
author: "Tommaso Rigon and Emanuele Aliverti"
bibliography: ["../biblio.bib"]
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Endometrial cancer study

*(This tutorial illustrates main results of the analysis in an easy-to-read format.
Complete code associated with this page is available in the file* 
[endometrial.Rmd](https://raw.githubusercontent.com/tommasorigon/logistic-bias-reduction/main/ENDOMETRIAL/endometrial.Rmd)),

The endometrial cancer study is illustrated in @Heinze2002, and focuses on $n = 79$ patients 
to evaluate the relationship between the histology of the endometrium (low against high), 
and three risk factors: neovasculation, pulsatility index of arteria uterina, and endometrium height. 
As shown in @Heinze2002, maximum likelihood for the coefficient $\beta_2$ (neovasculation) does not exist.

```{r endometrial,echo=F}
rm(list = ls())
library(knitr)
library(jtools)
```

```{r mlefit}
library(brglm2)
data(endometrial)
fit_mle <- glm(HG ~ NV + PI + EH, data = endometrial, family = binomial("logit"))
```

```{r mle,echo=F}
print(summ(fit_mle, model.info = F, model.fit = F,confint = F))
beta_mle <- coef(fit_mle)
sd_mle <- summary(fit_mle)$coefficients[, 2]
# I manually set these parameters to + Inf
beta_mle[2] <- sd_mle[2] <- Inf
```


## Regularization and bias correction 
As illustrated in Section 1 of the article, our approach relies on a Diaconis-Ylvisaker conjugate prior with carefully chosen hyperparameters. 
The resulting penalized likelihood can be expressed as a genuine Binomial likelihood, obtained replacing the Binomial observations via pseudo-counts. 
We construct such quantities in the variable `HG_DY`.

```{r data}
y <- endometrial$HG
X <- model.matrix(HG ~ NV + PI + EH, data = endometrial)
p <- ncol(X)
m <- nrow(X)
endometrial$HG_DY <- p / (p + m) * 0.5 + m / (p + m) * endometrial$HG
```

Our method can be estimated directly via `glm`, or other more efficient implementations for 
logistic regression models (refer to [high-dimensional](../TEST) for further examples in higher dimensions). 
Note that pseudo counts are not necessary integers, and this might prompt a warning when `glm` is called with `binomial` family.
Such warnings can be avoided relying on the `quasi` family with link and variance function corresponding
to binomial likelihood. In both cases, penalized likelihood optimization provide finite values for all coefficients.

```{r dy,warning=FALSE}
fit_dy <- glm(HG_DY ~ NV + PI + EH, data = endometrial, family = binomial("logit"))
# Avoid warnings
# fit_dy_now <- glm(HG_DY ~ NV + PI + EH, data = endometrial, family = quasi(link = "logit",variance = "mu(1-mu)"))
```

```{r dy-out,echo=F}
print(summ(fit_dy, model.info = F, model.fit = F,confint = F))
```

Table 1 in the paper compares the proposed approach with state-of-the art methods for 
bias-reduction, such as @Firth1993 and @Pagui2017, conveniently implemented in the 
R package `brglm2` [@brglm2]; refer also to @Kosmidis2020.
We also compare the proposed method with the approach of [@Clogg1991], which also relies on a 
conjugate prior specification inducing pseudo-counts and shrinking estimates toward the empirical 
proportion.

```{r bias correction,warning=FALSE}
# CLOGG ET AL. (1991)
endometrial$HG_CLOGG <- p / (p + m) * mean(endometrial$HG) + m / (p + m) * endometrial$HG
fit_clogg <- glm(HG_CLOGG ~ NV + PI + EH, data = endometrial, family = binomial("logit"))

# FIRTH (1993)
fit_firth <- glm(HG ~ NV + PI + EH,
  data = endometrial, family = binomial("logit"),
  method = "brglmFit", type = "AS_mean"
)

# KENNE PAGUI ET AL. (2017)
fit_kp <- glm(HG ~ NV + PI + EH,
  data = endometrial, family = binomial("logit"),
  method = "brglmFit", type = "AS_median"
)
```


```{r collect, echo = F}
beta_dy <- coef(fit_dy)
sd_dy <- summary(fit_dy)$coefficients[, 2]

beta_firth <- coef(fit_firth)
sd_firth <- summary(fit_firth)$coefficients[, 2]

beta_kp <- coef(fit_kp)
sd_kp <- summary(fit_kp)$coefficients[, 2]

beta_clogg <- coef(fit_clogg)
sd_clogg <- summary(fit_clogg)$coefficients[, 2]

estimates <- rbind(beta_mle, beta_dy, beta_clogg, beta_firth, beta_kp)
std_errors <- rbind(sd_mle, sd_dy, sd_clogg, sd_firth, sd_kp)
results_table = matrix(paste0(round(estimates,3)," (", round(std_errors,3),")"),ncol = 4)
row.names(results_table) = c("MLE", "Diaconis-Ylvisaker", "@Clogg1991", "@Firth1993", "@Pagui2017")
kable(results_table, digits = 3,format = "markdown", caption = "Table 1 of the paper: Estimated regression coefficients on the endometrial cancer study (and standard
errors)")
```

# References

